\vspace{-4em}Rethinking Model Architecture and Means of Delivery to get Video Interaction with Generative Neural Networks Online

### 2020-11-25 writing/paper/draft_02/draft_02.tex
Comp Graph diagrams

The Pix2Pix generator is an Autoencoder similar to the ``U-Net'' architecture\cite{1505.04597} but differs in some key ways, and we offer some further changes to drastically reduce the size. The generator consists of 8 2D convolution layers downsampling the input data in width and height, and 7 transpose convolutions upsampling the data back to original size. Each layer is batchnormalized\cite{1502.03167} apart from the first, and layers 8--10 have dropout\cite{JMLR:v15:srivastava14a} applied at a rate of 50\%. The Pix2Pix architecture includes no avg- or max-pooling as all filter kernels are 4$\times$4 and a stride of 2 is used. This has the effect of downsampling the layer input by half in width and height, thus circumventing the need for any kind of pooling, otherwise known as an \textit{all convolutional net}\cite{1412.6806}. The difference proposed in this work is to greatly reduce the number of filters used in each layer. The majority of trainable parameters are found in the middle of the generator as a result of the structure described. By reducing the number of filters in each layer by up to a factor of 8, very usable results were obtained whilst drastically reducing the model size --- more discussion on the impacts of a smaller model are found in Results.

\begin{small}
	\begin{table}
		\begin{center}
			\begin{tabular}{p{2cm}p{2cm}p{2cm}p{2cm}}
				\toprule
				Generator Layer & Pix2Pix & Our Small & Our Medium \\
				\midrule
				1               & 64      & 4         & 8          \\
				2               & 128     & 8         & 16         \\
				3               & 256     & 16        & 32         \\
				4               & 512     & 32        & 64         \\
				5               & 512     & 64        & 128        \\
				6               & 512     & 128       & 256        \\
				7               & 512     & 256       & 512        \\
				8               & 512     & 512       & 512        \\
				9               & 512     & 256       & 512        \\
				10              & 512     & 128       & 256        \\
				11              & 512     & 64        & 128        \\
				12              & 512     & 32        & 64         \\
				13              & 512     & 16        & 32         \\
				14              & 256     & 8         & 16         \\
				15              & 128     & 4         & 8          \\
				\bottomrule
			\end{tabular}
			\caption{Number of filters per layer in the generator. TODO: Replace table with graphic of model architecture, showing difference between model sizes.}
		\end{center}
	\end{table}
\end{small}

@@

\begin{comment}
It can be observed in the stack trace (using the developer tools of most modern browsers, in this case Firefox and Chrome were used) that TensorflowJS is still having to do a lot of the working moving data around and reshaping data in Javascript, while the heavy computation is deferred to shader programs. This moving data around comes at a great cost which is felt particualrly in the case of video based interaction. As a slight aside, a `pure' WebGL Autoencoder of the same stucture as the Pix2Pix generator was created simply to perform the same number of calculations in the same fashion to compare performance, and to give a glimpse into a novel approach to machine learning online which is perhaps more streamlined.
\end{comment}

\section{Results}
% 500-750 words
% -- 389 WORDS ---

@@

\begin{comment}
The aforementioned `pure' WebGL Autoencoder is a scalable CNN framework which holds the core functionality to perform the forward pass of an \textit{all convolutional} network, such as that of the Pix2Pix generator.
\end{comment}

@@

Given the smaller scale model with this workflow, the development process becomes faster and less daunting, allowing one to engage with the entire process more fully. Neural networks as a \textit{tool}, just like a paintbrush or a camera, is an exciting future in art. The iterative process involved in this work meant creating many datasets and training many models and assessing the outcome qualitatively, this helps develop a sense of \textit{how the model will react to a certain dataset}, and equally an \textit{intuition of the output}. The ability to iterate quickly is how, as a practitioner or an artist, one develops an intuition of ones tools and thus how one learns to control the medium --- a painter is not born with the ability to accurately mix color and control a brush, but it is learnt and honed over time.

This work is not only about creating art, or tools for artists. Democratising the process from start-to-finish is important to making the process accessible to all. We currently live with machine learning and artificial intelligence; as there appears to be no limits to the possibilities there is no doubt that it will find its way into even more aspects of our lives. Therefore it is imperative that the field of machine learning becomes more accessible and transparent. The first automobile was met with distrust[\emph{*find citation}] and seen as magic, the hope of this work is to dispel some myths of ML by giving the user their first ride in the car.

\begin{comment}
Given the remit of simplicity, an \textit{all \underline{WebGL} convolutional network} was created to mimic the structure of the Pix2Pix generator which shows a possible future of NN deployment on the web, specifically with video interaction in mind. All data is stored and processed in data structures suited to 2D textures which are quickly read and written to by GLSL shader programs.
\end{comment}

@@

\begin{comment}
\begin{appendices}
	\section{Training Process}
	\label{appendix:training}
	Some text that will be the appendix.

	\section{WebGL CNN}
	\label{appendix:webgl_cnn}
	Some text that will be the appendix.

\end{appendices}
\end{comment}

>>>>

The Pix2Pix generator is an Autoencoder similar to the ``U-Net'' architecture\cite{1505.04597} but differs in some key ways, and we offer some further changes to drastically reduce the size. The generator consists of 8 2D convolution layers downsampling the input data in width and height, and 7 transpose convolutions upsampling the data back to original size. Each layer is batchnormalized\cite{1502.03167} apart from the first, and layers 8--10 have dropout\cite{JMLR:v15:srivastava14a} applied at a rate of 50\%. The Pix2Pix architecture includes no avg- or max-pooling as all filter kernels are 4$\times$4 and a stride of 2 is used. This has the effect of downsampling the layer input by half in width and height, thus circumventing the need for any kind of pooling, otherwise known as an \textit{all convolutional net}\cite{1412.6806}. The difference proposed in this work is to greatly reduce the number of filters used in each layer. The majority of trainable parameters are found in the middle of the generator as a result of the structure described. By reducing the number of filters in each layer by up to a factor of 8, very usable results were obtained whilst drastically reducing the model size --- more discussion on the impacts of a smaller model are found in ~\ref{section:results}.

\begin{small}
	\begin{table}
		\begin{center}
			\begin{tabular}{p{2cm}p{2cm}p{2cm}p{2cm}}
				\toprule
				Generator Layer & Pix2Pix & Our Medium & Our Small \\
				\midrule
				1               & 64      & 8          & 4         \\
				2               & 128     & 16         & 8         \\
				3               & 256     & 32         & 16        \\
				4               & 512     & 64         & 32        \\
				5               & 512     & 128        & 64        \\
				6               & 512     & 256        & 128       \\
				7               & 512     & 512        & 256       \\
				8               & 512     & 512        & 512       \\
				9               & 512     & 512        & 256       \\
				10              & 512     & 256        & 128       \\
				11              & 512     & 128        & 64        \\
				12              & 512     & 64         & 32        \\
				13              & 256     & 32         & 16        \\
				14              & 128     & 16         & 8         \\
				15              & 64      & 8          & 4         \\
				\bottomrule
			\end{tabular}
			\caption{Number of filters per layer in the generator. TODO: Replace table with graphic of model architecture, showing difference between model sizes.}
		\end{center}
	\end{table}
\end{small}

\begin{figure}[h]
	\hspace*{-0.725in}
	\centering
	\begin{tabular}{cc}
		\subfloat{\includegraphics[width=3.2in]{medium_on_big}} & \subfloat{\includegraphics[width=3.2in]{small_on_medium}} \\
	\end{tabular}
	\caption{A visual comparison of our medium sized generator with the original Pix2Pix (left) and our small and medium generators (right).}
\end{figure}

@@

\section{Results}
\label{section:results}
% 500-750 words
% -- 389 WORDS ---

@@

@@

Given the smaller scale model with this workflow, the development process becomes faster and less daunting, allowing one to engage with the entire process more fully. Neural networks as a \textit{tool}, just like a paintbrush or a camera, is an exciting future in art. The iterative process involved in this work meant creating many datasets, training many models and assessing the outcome qualitatively, this helps develop a sense of \textit{how the model will react to a certain dataset}, and equally an \textit{intuition of the output}. The ability to iterate quickly is how, as a practitioner or an artist, one develops an intuition of ones tools and thus how one learns to control the medium --- a painter is not born with the ability to accurately mix color and control a brush, but it is learnt and honed over time.

This work is not only about creating art, or tools for artists. Democratising the process from start-to-finish is important to making the process accessible to all. We currently live with machine learning and artificial intelligence; as there appears to be no limits to the possibilities there is no doubt that it will find its way into even more aspects of our lives. Therefore it is imperative that the field of machine learning becomes more accessible and transparent to dispell any myths of magic.

@@
